Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network

### 2017-03-01 websci-2017/source/paper.tex
background

\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore}\cite{elmer-dewitt_first_1993}}

Much of the world's infrastructure is networked: power grids,
cellular networks, raods, and of course, the Internet.
With any critical infrastructure, the cost of failures can be
immense, so methods for tolerating various kind of faults within networks are an
important and ongoing area of research
\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
{\em Adversarial faults},
those in which an adversary can target attacks strategically,
deserve special attention.
Such attacks are both extremely difficult to guard against and 
often have important social implications.
In particular, censorship and surveillance are often achieved
by targeting central network locations and either blocking or capturing
the information flowing through them.
The Internet's decentralized design was motivated
by the need to withstand targeted attacks, such as nuclear strikes
\cite{baran_distributed_1964}.
But despite longstanding common wisdom \cite{elmer-dewitt_first_1993},
both theoretical results and recent events (described below)
have demonstrated that the Internet can be surprisingly vulnerable to attack.
Decentralization remains a promising approach towards
building resilient networks,
but there is a need to better understand the relationship between
decentralized network structure and adversarial fault tolerance.

The Internet's vulnerability to censorship and other targeted attacks
has been demonstrated by several recent events.
In 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
The action was initiated by government order in Pakistan,
and spread beyond Pakistan when a single misconfigured router allowed the false
routing information to propagate globally.
While the government order and router misconfiguration initiated the outage,
it was the structure of the Internet's router network that allowed a fault in a
single router to propagate globally.
And while the action was not an intentional attack against the global Internet,
the ability of an attacker to succeed without trying only strengthens our
claim of the Intenrnet's vulnerability.
In 2013, the Texas-based email provider Lavabit was compelled by to turn
their private SSL keys over to the FBI \cite{poulsen_edward_2013}.
In order to protect their users from surveillance,
Lavabit ceased operations.
While originally intended as surveillance,
this action effectively became an act of censorship.
In this case, the attack was again successful due to a highly centralized
architecture:
SSL keys under control of a single entity, in a single legal jurisdiction.
It is also important to note that there was no technological failure,
the attack was achieved through coercion.
So we see that such vulnerabilities are not limited to any one system or protocol,
but result from centralized structure itself.

@@

We also assume that nodes {\em trust} their immediate neighbors,
so Mal is unable to compromise Alice's node, or her direct neighbors.
In the commonly used {\em web of trust} approach
\cite{zimmermann_official_1995,ferguson_practical_2003},
we would extend that trust transitively to the entire network.
However, we make a weaker and more realistic assumption:
that trust is extended transitively to nodes within a fixed
distance.
So only the beginning and end of a path between Alice and Bob is trusted
against interference from Mal.
We call this assumption
{\em partial trust transitivity},
and refer to such paths as {\em partially trusted}.

@@

The reader may wonder whether it is realistic or useful to assume control over
the network structure.
We have already seen that whenever a single point of failure exists in the network,
there is a potential to exploit it through coercion,
without an attacker needing to compromise the technology.
So, {\em attack-tolerance cannot be achieved without the ability to influence
network structure}.
Luckily, there are scenarios in which network topology can be dictated.
Examples include overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
In general,
{\em when the need for attack-tolerance is high enough to warrant investment
in infrastructure, networks can be engineered and maintained as infrastructure}.
It is also worth noting that attack-tolerant networks may be sub-components of
larger, less-constrained systems.
For example, a single server might be replaced by a distributed network of servers,
each with different ownership, physical location, and legal jurisdiction,
without placing any unrealistic constraints on the clients connecting to
those servers.

We first describe how fault tolerance techniques can be adapted and
evaluated in a network setting with partial trust.
Generally, faults in network paths can be correlated,
preventing the appliation of standard fault tolerance techniques
\cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
which assume independent faults.
By using {\em independent paths},
which have no untrusted nodes in common,
we model communication across a complex network by a number of
redundant and independent channels.
Redundant messages can be sent across these channels in parallel,
a technique known as {\em concurrent multipath routing}
\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
and used for fault tolerance.
The receiver can then use the redundant messages to detect and/or correct
errors.
We formally evaluate the
effects of network structure on attack-tolerance and show that the probability
of an undetected error decreases exponentially with the number of
independent paths between source and destination,
even when no individual path is entirely trusted.

We also propose a novel concurrent multipath routing algorithm for the butterfly
topology.
The algorithm identifies independent paths,
which can combined with the fault-tolerant concurrent multipath routing
scheme above to
achieve a high level of adversarial fault tolerance on the butterfly topology.

Our main contributions are:
\begin{itemize}
\item{
We show that {\em independent paths} with {\em partial trust transitivity}
enable standard fault tolerance techniques to be applied to adversarial faults in
complex communication networks.
We find that the probability of detecting adversarial faults
approaches 1 exponentially as the number of independent paths between
sender and receiver increases;
}
\item{We present a scalable, efficient, and attack-tolerant concurrent
multipath routing algorithm on the butterfly network topology.}
\end{itemize}

This paper is organized as follows.
Section \ref{sec-related}. reviews background and related work.
Section \ref{sec-infrastructure}. discusses the desired properties of attack-tolerant network
infrastructure.
Section \ref{sec-ft}. describes adversarial fault tolerance on
structured networks.
Section \ref{sec-butterfly}. describes the concurrent multipath routing
algorithm for the butterfly network topology.
Section \ref{sec-discussion}. discusses the our results.
And Section \ref{sec-conclusion}. concludes.

@@

There has been considerable work on trust in network security.
Both centralized and decentralized approaches are commonly used to create
trust infrastructures.
Centralized approaches such as {\em public key infrastructure} (PKI)
suffer from a number of vulnerabilities
\cite{ellison_ten_2000},
which stem largely from the single points of failure inherent to
centralization.
The well-known and widely-used {\em web of trust} approach
\cite{zimmermann_official_1995,ferguson_practical_2003}
is a decentralized alternative.
In a web of trust,
individuals choose who they trust initially.
Trust is then extended to new individuals if they are vouched for by a
currently-trusted individual,
making it possible to quickly establish a large group of trusted nodes.
However, the web of trust approach assumes that trust is infinitely
transitive,
which is unrealistic
\cite{christianson_why_1997}.

Previous work on evaluating incorporating network structure into
trust models has focused on authentication
protocols, showing that independent paths can reduce an adversary's ability
to impersonate a target
\cite{levien_attack-resistant_2009}.
Other work has shown that identifying independent paths in arbitrary networks
is NP-hard and provided approximation algorithms
\cite{reiter_resilient_1998}.
Our work complements these by explicitly quantifying the effects of partial
trust transitivity,
and providing a routing algorithm to efficiently construct independent
paths on a structured network.

@@

There are relatively few attack-tolerance schemes
that focus on network structure,
compared to more popular cryptographic approaches
\cite{ferguson_practical_2003}.
Fiat and Saia described a scheme that combines the butterfly topology
with expander graphs to create a highly censorship-resistant,
content-addressable network \cite{fiat_censorship_2002},
although this scheme exhibits poor scaling due to a very high level of data
replication.
Perhaps the most mature structural solution is the Freenet collaboration
\cite{clarke_freenet:_2001}.
Freenet uses secret sharing
\cite{shamir_how_1979, blakley_safeguarding_1979}
and small-world routing
\cite{zhang_using_2002,kleinberg_small-world_2000}
to create a content-addressable network with a high level of both
confidentiality and censorship resistance.
Freenet guarantees that data is stored redundantly,
but still allows for centralized network structure,
and thus single points of failure,
as data travels from its origin to the redundant storage locations.

{\em Multipath routing} protocols send redundant information
over multiple paths when routing a message through a network,
in contrast to traditional {\em unipath} routing, which uses
a single path.
Multipath routing can have many benefits, including reduced congestion,
increased throughput, and more reliability
\cite{qadir_exploiting_2015}.
Many of these routing protocols offer increased confidentiality
\cite{zin_survey_2015}.
Some approaches utilize redundant paths as backups for increased
fault tolerance
\cite{alrajeh_secure_2013},
and some specifically protect against adversarial faults
\cite{kohno_improvement_2012, khalil_unmask:_2010, lou_h-spread:_2006}.
Most work on multipath routing has been motivated by applications related to
wireless sensor networks (WSNs),
and have thus focused on ad hoc, unstructured networks, often having a central
base station.
The method of Liu et al.
\cite{liu_secure_2012}
routes multiple messages first to random peers and then
to a central base station,
with the network edges constrained by sensors' physical location.
The butterfly routing algorithm we present takes a conceptually similar
approach in a different network setting.

>>>>

\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore} \cite{elmer-dewitt_first_1993}}

Much of the world's infrastructure is networked: power grids,
cellular networks, raods, and of course, the Internet.
As with any critical infrastructure, the cost of failures can be
immense, so methods for tolerating various kind of faults within networks are an
important and ongoing area of research
\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
{\em Adversarial faults},
those in which an adversary can target attacks strategically,
deserve special attention.
Such attacks are both extremely difficult to guard against and 
often have important social implications.
In particular, censorship and surveillance are often achieved
by targeting central network locations and either blocking or capturing
the information flowing through them.
The Internet's decentralized design was motivated
by the need to withstand targeted attacks, such as nuclear strikes
\cite{baran_distributed_1964}.
But despite longstanding common wisdom \cite{elmer-dewitt_first_1993},
both theoretical results and recent events (described below)
have demonstrated that the Internet can be surprisingly vulnerable to attack.
Decentralization remains a promising approach towards
building resilient networks,
but there is a need to better understand the relationship between
decentralized network structure and adversarial fault tolerance.

The Internet's vulnerability to censorship and other targeted attacks
has been demonstrated by several recent events.
In 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
The action was initiated by government order in Pakistan,
and spread beyond Pakistan when a single misconfigured router allowed the false
routing information to propagate globally.
While the government order and router misconfiguration initiated the outage,
it was the structure of the Internet's router network that allowed a fault in a
single router to propagate globally.
And while the action was not an intentional attack against the global Internet,
the ability of an attacker to succeed without even trying only highlights
the Internet's vulnerability to adversarial faults.

In 2013, the Texas-based email provider Lavabit was compelled by to turn
their private SSL keys over to the FBI \cite{poulsen_edward_2013}.
In order to protect their users from surveillance,
Lavabit ceased operations.
While originally intended as surveillance,
this action effectively became an act of censorship.
In this case, the attack was again successful due to a highly centralized
architecture:
SSL keys under control of a single entity, in a single legal jurisdiction.
It is also important to note that there was no technological failure,
the attack was achieved through coercion.
So we see that such vulnerabilities are not limited to any one system or protocol,
but result from centralized structure itself.

@@

We also assume that nodes {\em trust} their immediate neighbors.
So Mal is unable to compromise Alice's node, or her direct neighbors.
In the commonly used {\em web of trust} approach
\cite{zimmermann_official_1995,ferguson_practical_2003},
we would extend that trust transitively to the entire network.
However, we make a weaker and more realistic assumption:
that trust is extended transitively to nodes within a fixed
number of hops.
So only the beginning and end of a path between Alice and Bob is trusted
against interference from Mal.
We call this assumption
{\em partial trust transitivity},
and refer to such paths as {\em partially trusted}.

@@

The reader may wonder whether it is realistic or useful to assume control over
the network structure.
We have already seen that whenever a single point of failure exists in the network,
there is a potential for an attacker to exploit it through coercion,
without needing to compromise the technology.
So, {\em attack-tolerance cannot be achieved without the ability to influence
network structure}.
Luckily, there are scenarios in which network topology can be dictated.
Examples include overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
In general,
{\em when the need for attack-tolerance is high enough to warrant investment
in infrastructure, networks can be engineered and maintained as infrastructure}.
It is also worth noting that attack-tolerant networks may be sub-components of
larger, less-constrained systems.
For example, a single server might be replaced by a distributed network of servers,
each with different ownership, physical location, and legal jurisdiction,
without placing any unrealistic constraints on the clients connecting to
those servers.

We begin by describing how fault tolerance techniques can be adapted and
evaluated in a network setting with partial trust transitivity and
adversarial faults.
Generally, faults in network paths can be correlated,
preventing the appliation of standard fault tolerance techniques
\cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
which assume independent faults.
By constructing {\em independent paths},
which have no untrusted nodes in common,
we show how to model communication across a complex network in the presence
of correlated adversarial faults as communication across redundant
simple channels with random errors.
Redundant messages can be sent across these channels in parallel,
a technique known as {\em concurrent multipath routing}
\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
and used for fault tolerance.
The receiver can then use the redundant messages to detect and/or correct
errors.
We formally evaluate the
effects of network structure on attack-tolerance and show that the probability
of an undetected error decreases exponentially with the number of
independent paths between source and destination,
even when no individual path is entirely trusted.

We also propose a novel concurrent multipath routing algorithm for the butterfly
topology.
The algorithm constructs independent paths,
which can be combined with the fault-tolerant concurrent multipath routing
scheme above to
achieve a high level of adversarial fault tolerance on the butterfly topology.

Our main contributions are:
\begin{itemize}
\item{
We propose a novel method for extending standard fault tolerance techniques to
{\em adversarial} faults in {\em complex networks}.
We do so by modelling redundant independent paths
with partial trust transitivity as a single virtual channel,
and show that the probability of detecting adversarial faults
approaches 1 exponentially with the number of paths;
}
\item{We prove that the number of independent paths between two nodes
in a wraparound butterfly network with partial trust transitivity
increases exponentially with the trust radius;
}
\item{We present a scalable, efficient, and attack-tolerant concurrent
multipath routing algorithm on the butterfly network topology.}
\end{itemize}

This paper is organized as follows.
Section \ref{sec-related} reviews background and related work.
Section \ref{sec-ft} describes adversarial fault tolerance on
structured networks.
Section \ref{sec-butterfly} describes the concurrent multipath routing
algorithm for the butterfly network topology
Section \ref{sec-discussion} discusses the our results.
And Section \ref{sec-conclusion} concludes.

@@

There has been considerable work on trust in network security.
Both centralized and decentralized approaches are commonly used to create
trust infrastructures.
Centralized approaches such as {\em public key infrastructure} (PKI)
suffer from a number of vulnerabilities
\cite{ellison_ten_2000},
which stem largely from the single points of failure inherent to
centralization.
The well-known and widely-used {\em web of trust} approach
\cite{zimmermann_official_1995,ferguson_practical_2003}
is a decentralized alternative.
In a web of trust,
individuals choose who they trust initially.
Trust is then extended to new individuals if they are vouched for by a
currently-trusted individual,
making it possible to quickly establish a large group of trusted nodes.
However, web of trust's assumption of infinite transitivity is unrealistic
\cite{christianson_why_1997},
and does not distinguish between paths of different lengths.
Our work addresses both of these limitations by incorporating a more realistic
assumption of partial transitivity.

Previous work on evaluating incorporating network structure into
trust models has focused on authentication
protocols, showing that independent paths can reduce an adversary's ability
to impersonate a target
\cite{levien_attack-resistant_2009}.
Other work has shown that identifying independent paths in arbitrary networks
is NP-hard and provided approximation algorithms
\cite{reiter_resilient_1998}.
Our work complements these by introducing the partial trust assumption and
explicitly quantifying its relationship to network structure and adversarial
fault tolerance.
As an alternative to the NP-hard task of finding independent paths on arbitrary
networks, we provide a routing algorithm to efficiently construct independent
paths on a the butterfly topology.

@@

There are relatively few attack-tolerance schemes
that focus on network structure,
compared to more popular cryptographic approaches
\cite{ferguson_practical_2003}.
All existing attack-tolerant networks we are aware of are content-addressable
networks: data is routed to and from storage nodes rather than between sender
and receiver.
Fiat and Saia described a scheme that combines the butterfly topology
with expander graphs to create a highly censorship-resistant,
content-addressable network \cite{fiat_censorship_2002},
although this scheme does not scale well and is impractical due to a
high level of data replication.
Perhaps the most mature structural solution is the Freenet collaboration
\cite{clarke_freenet:_2001}.
Freenet uses secret sharing
\cite{shamir_how_1979, blakley_safeguarding_1979}
and small-world routing
\cite{zhang_using_2002,kleinberg_small-world_2000}
to create a content-addressable network with a high level of both
confidentiality and censorship resistance.
Freenet guarantees that data is stored redundantly,
but still allows for centralized network structure,
and thus single points of failure,
as data travels from its origin to the redundant storage locations.
Unlike the above content-addressable networks, our proposal is purely network based
and does not require nodes to store data indefinitely.
Our proposal also improves on the scalablity of Fiat and Saia's work,
and does not rely on assumptions about existing social network structure.

send redundant information
over multiple paths when routing a message through a network,

{\em Multipath routing} protocols identify multiple paths between
source and destination
in contrast to traditional {\em unipath} routing, which uses
a single path.
The special case of {\em concurrent} multipath routing uses mutliple paths
simultaneously.
Multipath routing has many applications, including reduced congestion,
increased throughput, and more reliability
\cite{qadir_exploiting_2015}.
Many of these routing protocols offer increased confidentiality
\cite{zin_survey_2015}.
Some approaches utilize redundant paths as backups for increased
fault tolerance
\cite{alrajeh_secure_2013},
and some specifically protect against adversarial faults
\cite{kohno_improvement_2012, khalil_unmask:_2010, lou_h-spread:_2006}.
Most work on multipath routing has been motivated by applications related to
wireless sensor networks (WSNs),
and have thus focused on ad hoc, unstructured networks, often having a central
base station.
The method of Liu et al.
\cite{liu_secure_2012}
routes multiple messages first to random peers and then
to a central base station,
with the network edges constrained by sensors' physical location.
We have found only few examples in the existing literature of applications of
concurrent multipath routing to {\em adversarial} fault tolerance,
and all have focused on ad-hoc wireless sensor networks, without attention
to the role of network structure.
The alogorithm we present for the butterfly topology complements existing work
by addressing cases where links are not constrained by physical
distance,
and where network structure can be engineered for greater attack-tolerance.
